A thesis draft

### 2018-11-19 res/sections/08-Background.tex
Add part on NVFI and NFV challenges

\subsection{Network Virtualization}
%http://www.csc.villanova.edu/~nadi/csc8580/S11/DeeptiNune.pdf
%https://www.sciencedirect.com/science/article/abs/pii/S1389128609003387
Virtualization, in the networking context, refers to the act of combining
hardware, software and functionalities into a virtual network. This abstraction
runs on the top of a physical network in a hyper-visor, decoupling the logical
network from the hardware one. Network virtualization allow multiple,
heterogeneous networks to cohabit, increasing the flexibility and manageability.
The idea of network virtualization came from the decoupling of the role of the
traditional ISP into two different entities: who provides the infrastructure
(InPs) and who provides the services(SPs). In that context it is possible to
aggregate resources, even from different InPs, creating a virtual network,
offering an end-to-end service~\cite{chowdhury2009network}. Most known
instances of NV are \emph{VPNs}, dedicated networks that allows the
communication among multiple sites, using
private and secure tunnels, relying on the public network such as Internet.
Other examples of network virtualization are \emph{overlay networks}. They are
logical networks, build on the top of an existing physical network, typically
implemented at the application layer.

\subsection{Network Function Virtualization}
The \emph{Network Function Virtualization} (NFV) concept refers to the use of 
virtualization technologies to virtualize network nodes functions. VNFs may run 
different processes and software, on the top of servers, switches, storage 
devices or on cloud computing infrastructures, that substitute custom hardware. 
That approach worked well in the past, because of the need to follow rigorous 
standards, providing at the same time high quality, but it led to long product 
cycles and high specialized proprietary hardware. New Internet scenarios, that 
include IoT and the increasing number of services to spread multimedia data 
(Skype, Netflix, Spotify, \dots), requires more flexible functions and faster 
network deployments. The NFV technology is exploited in a framework, organized
at least by three components:
\begin{enumerate}
 \item a software that performs the network function that must be deployed,
 that is the VNF;
 \item NFV Infrastructure (NFVI) that is composed by all hardware and software 
components that builds the environment where the VNFs are deployed;
 \item NFV Management And Orchestration (NFV-MANO) architectural framework that
 is all the functional blocks, data repositories used by these blocks,
 reference points and interfaces through which these functional blocks exchange
 information for the purpose of managing and orchestrating NFVI and VNFs.
\end{enumerate}

>>>>

\subsection{Network Virtualization}
Virtualization, in the networking context, refers to the act of combining
hardware, software and functionalities into a virtual network. This abstraction
runs on the top of a physical network in a hyper-visor, decoupling the logical
network from the hardware one. Network virtualization allow multiple,
heterogeneous networks to cohabit, increasing the flexibility and manageability.
The idea of network virtualization came from the decoupling of the role of the
traditional ISP into two different entities: who provides the infrastructure
(InPs) and who provides the services(SPs). In that context it is possible to
aggregate resources, even from different InPs, creating a virtual network,
offering an end-to-end service~\cite{chowdhury2009network}. Most known
instances of NV are \emph{VPNs}, dedicated networks that allows the
communication among multiple sites, using
private and secure tunnels, relying on the public network such as Internet.
Other examples of network virtualization are \emph{overlay networks}. They are
logical networks, build on the top of an existing physical network, typically
implemented at the application layer.

\subsection{Network Function Virtualization}
The \emph{Network Function Virtualization} (NFV) concept refers to the use of 
virtualization technologies to virtualize network nodes functions. This
emerging technology focuses on the decoupling of function logic and hardware,
allowing VNFs to run on the top of servers, switches, storage 
devices or on cloud computing infrastructures. In other word the software logic
is separated from the custom platform on which it runs, enabling the use of
commodity hardware~\cite{gray2016network}.
That approach worked well in the past, because of the need to follow rigorous 
standards, providing at the same time high quality, but it led to long product 
cycles and high specialized services that use a combination of software and
hardware resources, typically vendor-specific, and only largely
non-interoperable within other vendor solutions. New Internet scenarios, that 
include IoT and the increasing number of services to spread multimedia data 
(Skype, Netflix, Spotify, \dots), requires more flexible functions and faster 
network deployments. 
This detachment provides greater flexibility to scale VNF based on traffic and
performances and also improve the automation of function deploy and update:
removing dedicated hardware also will reduce costs machines and for
maintenance.

The NFV technology is exploited in a framework, organized
at least by three components~\cite{etsi2013gs}, as represented in
Figure~\ref{chap:background:img:etsi_arch}:
\begin{enumerate}
  \item a software, able to run on over the NFVI, that performs the network
  function that must be deployed, that is the VNF;
  \item NFV Infrastructure (NFVI) that is composed by all physical resources
  and how these can be virtualized, that forms the environment where the VNFs
  are deployed;
  \item NFV Management And Orchestration (NFV-MANO) that covers the
  orchestration and management of both physical and logical resources, the
  functional blocks, data repositories used by these blocks, reference points
  and interfaces through which these functional blocks exchange
  information for the purpose of managing and orchestrating NFVI and VNFs.
\end{enumerate}

\begin{figure}
  \centering
  \includegraphics[scale=2]{etsi_arch}
  \caption{Network Function Virtualization reference
  architecture~\cite{etsi2013gs}}
  \label{chap:background:img:etsi_arch}
\end{figure}

\subsection{Network Function Virtualized Infrastructure}
As mentioned before, NFVI is the combination of physical and logical resources,
that combined together create the environment on which the VNFs are deployed
and run. Physical supplies can be proprietary systems, commercial-off-the-shelf
hardware, software and network resources or cloud computing capabilities. The
abstraction of resources is made taking advantage of a virtualization layer,
that hides the underlying infrastructure. Storage and computing resources may
be represented in terms of virtual machines or
containers~\cite{mijumbi2016network}.

\subsection{VNF challenges}
This new concept open a plethora of challenges~\cite{han2015network}. Regarding
performances, functions running on general purposes servers cloud affect
throughput and latency, moreover the capacity of the virtualized function can
be less than the hardware and software version. To reach comparable levels in
terms of performances efficient algorithms are needed, as a excellent resource
management. Another problems is the VNF management. In case of hardware
functions resources are almost equivalent, at the contrary, virtualized ones
could have costs and values that differs significantly depending on the point
of presence on which are deployed. On the other hand, since VNFs are software
components, can elastically scale up and down, depending on the traffic loads
and user demand. Another management challenge refers to keeping track where the
VNF is running, that is of paramount importance in case of issues. Related to
possible dysfunctions, software function must reach a certain level of
reliability. Generally, purpose-build network equipment can ensure the
five-nines reliability requirement, that is more challenging to follow in that
scenario due to the higher number of possible points of failure, such as during
the scaling operation or VNF migrating from a point of presence to another.
Finally even security must be taken into account. In fact, infrastructure on
which VNFs run can be outsourced to third parties. Furthermore, orchestration
and management components must implement some sort of intrusion prevention and
detection system and the sharing of resources can introduce security threads.
